### 2020-06-05 4_metg.tex
Define METTS.

METG has a well-defined
relationship with quantities of interest such as
weak and strong scaling. Figures~\ref{fig:weak-scaling-mpi} and
\ref{fig:strong-scaling-mpi} show the weak and strong scaling of MPI Task Bench running a stencil pattern at a variety of problem sizes. In these
figures, the vertical axis is shown as wall time to emphasize the
relationship to time-to-solution, but it could equivalently be shown
as task granularity (as the number of tasks per execution is
fixed). Intuitively, at
larger problem sizes MPI is perfectly efficient. This can be seen at
the top of each figure, with flat lines when weak scaling and
ideally-sloped downward lines when strong scaling. Inefficiency begins
to appear at smaller problem sizes, towards the bottom of the graph,
where lines become more compressed. At the
very bottom, the lines compress together as running time becomes
dominated by overhead. Note that the contour of the bottom of each
graph is identical and conforms to the METG curve (marked by the red,
dashed line).

>>>>

METG has a well-defined
relationship with quantities of interest such as
time to solution. Figures~\ref{fig:weak-scaling-mpi} and
\ref{fig:strong-scaling-mpi} show the weak and strong scaling of MPI
Task Bench running a stencil pattern at a variety of problem
sizes. {\color{blue} To show the relationship with METG, the red,
  dashed line in each graph marks the \emph{minimum effective time to
    solution} (METTS). This is computed by multiplying METG by the
  number of tasks per core (in this case 1000). METG is computed
  separately for each node count, so METTS varies at scale.}
Intuitively, at
larger problem sizes MPI is perfectly efficient. This can be seen at
the top of each figure, with flat lines when weak scaling and
ideally-sloped downward lines when strong scaling. Inefficiency begins
to appear at smaller problem sizes, towards the bottom of the graph,
where lines become more compressed. At the
very bottom, the lines compress together as running time becomes
dominated by overhead. Note that the contour of the bottom of each
graph is identical and conforms to the METTS curve.

### 2020-06-05 5_evaluation.tex
Define METTS.

We can use METG to predict the scalability of a code. Figure
\ref{fig:strong-scaling} shows strong scaling for three Task Bench
implementations with the stencil pattern. Lines marked ``actual''
represent strong scaling measurements, while ones marked ``limit'' are
computed by multiplying METG(50\%) by tasks per core (in this case,
1000). Intuitively, ``limit'' is the smallest time to solution that
can be achieved for \emph{any} problem size at that node count, while
maintaining 50\% efficiency. Note that ``actual'' eventually falls
below ``limit'' because ``actual'' is not constrained to 50\%
efficiency and eventually drops below this level.

There are two points of interest on each ``limit'' curve: the point
where it intersects ``actual'' and the point where it intersects an
ideal scaling curve, computed by taking the initial time to solution
and assuming linear scaling. These points are marked in
Figure~\ref{fig:strong-scaling} with a black square and black circle,
respectively. Notably, the intersection between ideal and ``limit''
requires only a single node run, along with measurements of METG(50\%)
at each node count. Using this data it is possible to predict the
strong scalability of the code: i.e., the smallest time to solution,
and the number of nodes at which that time is achieved, while
maintaining at least 50\% efficiency. In
Figure~\ref{fig:strong-scaling} this can be seen as the distance
between the black square and circle; these are separated by a factor of
$1.22\times$ in node count, and $1.27\times$ in time to solution.

Table~\ref{tab:metg-predict-strong} expands this comparison to the 4
patterns tested in Figure~\ref{fig:metg-compute} for all 11
programming systems that scale well enough to evaluate the separation
between the limit-ideal and limit-actual intersections. We see that
overall, the mean separation is less than $2.15\times$ in node count
and less than $1.31\times$ in time to solution, making this an
accurate way to predict strong scaling in 4 patterns we tested in
Section~\ref{subsec:scalability}.

>>>>

We can use METG to predict scalability. Figure
\ref{fig:strong-scaling} shows three
implementations strong scaling the stencil pattern. Lines marked ``actual''
represent strong scaling measurements. The METTS lines are calculated as described in Section~\ref{sec:metg}, by multiplying multiplying METG(50\%) by tasks per core (1000). Intuitively, METTS is the smallest time to solution that
can be achieved for \emph{any} problem size at that node count, while
maintaining 50\% efficiency. Note that ``actual'' eventually falls
below METTS because ``actual'' is not constrained to 50\%
efficiency and eventually drops below this level.

There are two points of interest on each METTS curve: the point
where it intersects ``actual'' and the point where it intersects an
ideal scaling curve, computed by taking the initial time to solution
and assuming linear scaling. These points are marked in
Figure~\ref{fig:strong-scaling} with a black square and black circle,
respectively. Notably, the intersection between ideal and METTS
requires only a single node run, along with measurements of METG(50\%)
at each node count. Using this data it is possible to predict the
strong scalability of the code: i.e., the smallest time to solution,
and the number of nodes at which that time is achieved, while
maintaining at least 50\% efficiency. In
Figure~\ref{fig:strong-scaling} this can be seen as the distance
between the black square and circle; these are separated by a factor of
$1.22\times$ in node count, and $1.27\times$ in time to solution.

Table~\ref{tab:metg-predict-strong} expands this comparison to the 4
patterns tested in Section~\ref{subsec:scalability} for all 11
programming systems that scale well enough to evaluate the separation
between the METTS-ideal and METTS-actual intersections. We see that
overall, the mean separation is at most $2.15\times$ in node count
and at most $1.31\times$ in time to solution, making this an
accurate way to predict strong scaling in the 4 patterns we tested in
Section~\ref{subsec:scalability}.

### 2020-06-05 e11_metg_predict_strong.tex
Minor fix.
Define METTS.

\begin{table}[t]
\color{blue}
\centering
\small
\begin{tabular}{r | c | c | c | c}
        & \multicolumn{2}{|c|}{Nodes} & \multicolumn{2}{|c}{Time to Solution} \\
Pattern & mean & std. dev. & mean & std. dev. \\
\hline
stencil & $1.61\times$ & 0.324 & $1.29\times$ & 0.238 \\
nearest & $1.77\times$ & 0.253 & $1.17\times$ & 0.341 \\
spread  & $1.39\times$ & 0.322 & $1.28\times$ & 0.207 \\
nearest, 4 graphs & $2.15\times$ & 0.635 & $1.31\times$ & 0.777
\end{tabular}

\vspace{-0.20cm}
\caption{\color{blue} Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Figure~\ref{fig:metg-compute}.\label{tab:metg-predict-strong}}
\vspace{-0.5cm}
\end{table}

>>>>

\begin{table}[t]
\color{blue}
\centering
\small
\begin{tabular}{r | c | c | c | c}
        & \multicolumn{2}{|c|}{Nodes} & \multicolumn{2}{|c}{Time to Solution} \\
Pattern & mean & std.~dev. & mean & std.~dev. \\
\hline
stencil & $1.61\times$ & 0.324 & $1.29\times$ & 0.238 \\
nearest & $1.77\times$ & 0.253 & $1.17\times$ & 0.341 \\
spread  & $1.39\times$ & 0.322 & $1.28\times$ & 0.207 \\
nearest, 4 graphs & $2.15\times$ & 0.635 & $1.31\times$ & 0.777
\end{tabular}

\vspace{-0.20cm}
\caption{\color{blue} Average factor of separation between METTS-ideal and METTS-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
\vspace{-0.5cm}
\end{table}
